---
title: "9.1"
author: "em"
date: "July 30, 2016"

output:
  html_document:
    keep_md: yes
---

Do a Bayesian analysis of hypocotyl length (hyp) in the attached data sheet.

1) Consider treatment effects (trt) species effects (species) and their interaction.  What is the best model given these possible predictors?

First with treatment only
```{r}
setwd("~/Documents/Rclub-rethinking_Emily.Josephs/Assignment_Chapter_09")
d = read.csv('TomatoR2CSHL.csv', header=T)
d$trtNum = sapply(d$trt, function(x){if(x == 'H'){return(1)} else {return(-1)}})
d.trim = d[,c('trtNum','hyp')]
library(rethinking)

#trt only
m1 <- map2stan(
  alist(
    hyp~dnorm(mu,sigma),
    mu <- a + bT*trtNum,
    a ~ dnorm(10,10),
    bT ~ dnorm(0,1),
    sigma ~ dcauchy(0,2)),
  data=d.trim
     )

plot(m1)
precis(m1)
```
So with treatment only, there's a negative relationship where the high treatment has shorter hypocotyls. Is this what the data looks like?
```{r}

plot(jitter(d$trtNum), d$hyp, xaxt = "n", xlim = c(-2,2), xlab = "")
axis(1, at = c(-1,1), labels = c("low","high"))

```
Yep

Species only. I'm not sure how to deal with the categorical variables here, but I copied Julin's approach.

```{r}
library(reshape2)
data.species <- d[,c("hyp","species")]
data.species$id <- 1:nrow(data.species)
data.species <- dcast(data.species, hyp + id ~ species, value.var="species", fun.aggregate = length)
colnames(data.species) <- sub(". ","_",fixed = TRUE, colnames(data.species))

#species only

m2 <- map2stan(
  alist(
    hyp~dnorm(mu,sigma),
    mu <- a + bCh*S_chilense + bCm*S_chmielewskii + bH*S_habrochaites + bPn*S_pennellii + bPr*S_peruvianum,
    a ~ dnorm(10,10),
    c(bCh, bCm, bH, bPn, bPr) ~ dnorm(0,10),
    sigma ~ dcauchy(0,1)),
  data=data.species, chains=4
     )

plot(m2)
precis(m2)
plot(precis(m2))

```
I'm not totall sure how to interpret this -- essentially, all of the species beta estimates overlap 0 except for Peruvianum. Does that mean that only peruvianum shows a species specific effect?

```{r}
plot(d$species, d$hyp, ylab = "hypocotyl")
a1 <- aov(d$hyp ~ d$species)
summary(a1)
p1 <- TukeyHSD(x=a1, 'd$species', conf.level=0.95)
p1

```



2) Use "S. chmielewskii" as the reference.  For each of the other species evaluate the hypothesis that their response to shade differs from S. chmielewskii.



