---
title: "chapter7"
author: "em"
date: "June 5, 2016"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rethinking)
set.seed(10)
```

##7E1

1) Temperature

2) Field of study

3) Is the car broken?

##7E2
3

##7E3
1) Onion caramelization ~ Bt * temperature + Bw * moisture

2) Car speed ~ Bc * cylinders + Bi * fuel injector goodness

3) Political beliefs ~ Gp * parents' beliefs + Bf * friends' beliefs
Gp = Bp + Bpf * friends' beliefs
The Bpf term specifies whether the friends' beliefs matter or not.

4) Intelligence ~ Bs * socialness + Ba * appendages

##7M1

```{r, include=FALSE, eval=FALSE}
library(rethinking)
data(tulips)
d <- tulips
str(d)
m7.6 <- map(
  alist(
    blooms ~ dnorm(mu,sigma),
    mu <- a + bW*water + bS*shade,
    a ~ dnorm(0,100),
    bW ~ dnorm(0,100),
    bS ~ dnorm(0,100),
    sigma ~ dunif(0,100)
  ), data=d, method="Nelder-Mead", control=list(maxit=1e4))

m7.7 <- map(
  alist(
    blooms ~ dnorm(mu,sigma),
    mu <- a + bW*water + bS*shade + bWS*water*shade,
    a ~ dnorm(0,100),
    bW ~ dnorm(0,100),
    bS ~ dnorm(0,100),
    bWS ~ dnorm(0,100),
    sigma ~ dunif(0,100)
  ), data=d, method="Nelder-Mead", control=list(maxit=1e4))

coeftab(m7.6, m7.7)  
precis(m7.6)
precis(m7.7)
compare(m7.6, m7.7)

d$shade.c <- d$shade - mean(d$shade)
d$water.c <- d$water - mean(d$water)


m7.8 <- map(
  alist(
    blooms ~ dnorm(mu,sigma),
    mu <- a + bW*water.c + bS*shade.c,
    a ~ dnorm(0,100),
    bW ~ dnorm(0,100),
    bS ~ dnorm(0,100),
    sigma ~ dunif(0,100)
  ), data=d, start=list(a=mean(d$blooms), bW=0, bS = 0, sigma=sd(d$blooms))
  )

m7.9 <- map(
  alist(
    blooms ~ dnorm(mu,sigma),
    mu <- a + bW*water.c + bS*shade.c + bWS*water.c*shade.c,
    a ~ dnorm(0,100),
    bW ~ dnorm(0,100),
    bS ~ dnorm(0,100),
    bWS ~ dnorm(0,100),
    sigma ~ dunif(0,100)
  ), data=d, start=list(a=mean(d$blooms), bW=0, bS = 0, sigma=sd(d$blooms)))

coeftab(m7.8, m7.9)
precis(m7.9)
#bWS is negative because water and shade have opposite effects on blooms

par(mfrow=c(1,3))

shade.seq <- -1:1
for (w in -1:1){
  dt <- d[d$water.c == w,]
  plot(blooms ~ shade.c, data=dt, col=rangi2, main=paste("water.c =",w), xaxp = c(-1,1,2), ylim=c(0,362), xlab = "shade (centered)")
  mu <- link(m7.9, data=data.frame(water.c=w, shade.c=shade.seq))
  mu.mean <- apply(mu,2,mean)
  mu.PI <- apply(mu, 2, PI, prob = 0.97)
  lines(shade.seq, mu.mean)
  lines(shade.seq, mu.PI[1,], lty=2)
  lines(shade.seq, mu.PI[2,], lty=2)
  }

```

There is clearly an interaction between temperature, water, and shade.


##7M2

```{r, eval=FALSE}
blooms ~ a + bW*water + bS*shade + bWS*water*shade + bWST*temp*water*shade 

When temp is cool, temp=0. When temp is warm, temp = 1 and bWST*water*shade is equal to the sum of the other terms times 1.
```


##7M3

So, I think this biological interaction could be linear if raven population = wolf population + amount of food. However, the way it's described it seems like there's some sort of multiplicative thing happening since if there are no wolves, there are no ravens, such that raven population = wolf population * amount of food.


```{r, eval=F, include=F}
#wolfCount = round(rnorm(100)*50)+150
#hist(wolfCount, col="darkgray", border="white", breaks=20)
wolfCount = rnorm(100)
foodQuality = runif(100,0,1)
hist(foodQuality, col="darkgray", border="white", breaks=20)
ravenCount = wolfCount*foodQuality
hist(ravenCount, col="darkgray", border="white", breaks=20)

d = data.frame(wolfCount,foodQuality,ravenCount)

mr <- map(
  alist(
    ravenCount ~ dnorm(mu,sigma),
    mu <- a + bF*foodQuality + bW*wolfCount + bWF*foodQuality*wolfCount,
    a ~ dnorm(0,100),
    bF ~ dnorm(0,100),
    bW ~ dnorm(0,100),
    bWF ~ dnorm(0,100),
    sigma ~ dunif(0,100)
  ), data=d,  method="Nelder-Mead", control=list(maxit=1e4))

coeftab(mr)
precis(mr)

```

```
##7H1

```{r}
data(tulips)
d <- tulips

d$bed_id <- coerce_index(d$bed)
d$shade.c <- d$shade - mean(d$shade)
d$water.c <- d$water - mean(d$water)

m7.10 <- map(
  alist(
    blooms ~ dnorm(mu,sigma),
    mu <- a + bW*water.c + bS*shade.c + bWS*water.c*shade.c + bB*bed_id,
    a ~ dnorm(0,100),
    bW ~ dnorm(0,100),
    bS ~ dnorm(0,100),
    bB ~ dnorm(0,100),
    bWS ~ dnorm(0,100),
    sigma ~ dunif(0,100)
  ), data=d, start=list(a=mean(d$blooms), bW=0, bS = 0, bB = 0,sigma=sd(d$blooms)))

precis(m7.10)

```

##7H2


```{r}
m7.9 <- map(
  alist(
    blooms ~ dnorm(mu,sigma),
    mu <- a + bW*water.c + bS*shade.c + bWS*water.c*shade.c,
    a ~ dnorm(0,100),
    bW ~ dnorm(0,100),
    bS ~ dnorm(0,100),
    bWS ~ dnorm(0,100),
    sigma ~ dunif(0,100)
  ), data=d, start=list(a=mean(d$blooms), bW=0, bS = 0, sigma=sd(d$blooms)))

coeftab(m7.9, m7.10)

compare(m7.9, m7.10)
```

Bed has a positive affect on the number of blooms, but it doesn't seem like this affects the coefficients for any of the other variables.

##7H3

```{r}
data(rugged)
d <- rugged[ complete.cases(rugged$rgdppc_2000) , ]
d$loggdp = log(d$rgdppc_2000)

m.full <- map(
alist(
loggdp ~ dnorm(mu,sigma),
mu <- a + bR*rugged + bAR*rugged*cont_africa + bA*cont_africa,
a ~ dnorm(0,100),
bA ~ dnorm(0,100),
bR ~ dnorm(0,100),
bAR ~ dnorm(0,100),
sigma ~ dunif(0,100)
),data=d,  start=list(a=mean(d$loggdp), bA=0, bR = 0, bAR=0,sigma=sd(d$loggdp)))

precis(m.full)

noSey = d[d$country != "Seychelles",]

m.noSey <- map(
alist(
loggdp ~ dnorm(mu,sigma),
mu <- a + bR*rugged + bAR*rugged*cont_africa + bA*cont_africa,
a ~ dnorm(0,100),
bA ~ dnorm(0,100),
bR ~ dnorm(0,100),
bAR ~ dnorm(0,100),
sigma ~ dunif(0,100)
),data=noSey,  start=list(a=mean(d$loggdp), bA=0, bR = 0, bAR=0,sigma=sd(d$loggdp)))

precis(m.noSey)

coeftab(m.full, m.noSey)
```

a) It still seems like the effect of ruggedness depends on continent and the relationship hasn't changed that much.


```{r}
par(mfrow=c(2,2))
rugged.seq = seq(0,7, by = 0.5)

for ( w in 0:1 ) {
dt <- d[d$cont_africa==w,]
plot( loggdp ~ rugged , data=dt , col=rangi2 ,
main=paste("cont_africa =",w, ", all data") , ylim=c(5,12) , xlim = c(0,7),xlab="ruggedness" )
mu <- link( m.full , data=data.frame(cont_africa=w,rugged=rugged.seq) )
mu.mean <- apply( mu , 2 , mean )
mu.PI <- apply( mu , 2 , PI , prob=0.97 )
lines( rugged.seq , mu.mean )
lines( rugged.seq , mu.PI[1,] , lty=2 )
lines( rugged.seq , mu.PI[2,] , lty=2 )
}

for ( w in 0:1 ) {
dt <- d[noSey$cont_africa==w,]
plot( loggdp ~ rugged , data=dt , col=rangi2 ,
main=paste("cont_africa =",w, ", no Seychelles") , ylim=c(5,12) , xlim = c(0,7),xlab="ruggedness" )
mu <- link( m.noSey , data=data.frame(cont_africa=w,rugged=rugged.seq) )
mu.mean <- apply( mu , 2 , mean )
mu.PI <- apply( mu , 2 , PI , prob=0.97 )
lines( rugged.seq , mu.mean )
lines( rugged.seq , mu.PI[1,] , lty=2 )
lines( rugged.seq , mu.PI[2,] , lty=2 )
}
```

b) the effect of ruggedness still depends on continent, but the affect in Africa seems a bit weaker.

```{r}
m.noSey.1 <- map(
alist(
loggdp ~ dnorm(mu,sigma),
mu <- a + bR*rugged,
a ~ dnorm(0,100),
bR ~ dnorm(0,100),
sigma ~ dunif(0,100)
),data=noSey,  start=list(a=mean(d$loggdp), bR = 0, sigma=sd(d$loggdp)))

m.noSey.2 <- map(
alist(
loggdp ~ dnorm(mu,sigma),
mu <- a + bR*rugged + bA*cont_africa,
a ~ dnorm(0,100),
bA ~ dnorm(0,100),
bR ~ dnorm(0,100),
sigma ~ dunif(0,100)
),data=noSey,  start=list(a=mean(d$loggdp), bA=0, bR = 0, sigma=sd(d$loggdp)))


m.noSey.3 <- map(
alist(
loggdp ~ dnorm(mu,sigma),
mu <- a + bR*rugged + bAR*rugged*cont_africa + bA*cont_africa,
a ~ dnorm(0,100),
bA ~ dnorm(0,100),
bR ~ dnorm(0,100),
bAR ~ dnorm(0,100),
sigma ~ dunif(0,100)
),data=noSey,  start=list(a=mean(d$loggdp), bA=0, bR = 0, bAR=0,sigma=sd(d$loggdp)))

compare(m.noSey.1, m.noSey.2, m.noSey.3)

```

##7H4


```{r}